The Effect of Racial and Ethnic Attitudes on Asian Identity in the U.S

### 2024-09-18 my_paper/manuscript_online_appendix.tex
write intro

\documentclass[12pt,english]{article}

\input{fragments/preamble_packages.tex}
\input{fragments/preamble_layout.tex}
\input{fragments/preamble_citation.tex}
\bibliography{bib/AttitudesPaper}

\makeatother

\begin{document}

\title{The Effect of Racial and Ethnic Attitudes on Asian Identity in the U.S}

\author{\href{https://hussainhadah.com/}{Hussain Hadah} \thanks{Hussain Hadah: Department of Economics, Tulane University, 6823 St. Charles Ave., New Orleans, LA 70118, United States (e-mail: \href{mailto:hhadah@tulane.edu}{hhadah@tulane.edu}). I thank Professors Willa Friedman, Chinhui Juhn, Vikram Maheshri, and Yona Rubinstein for their support and advice. I also thank Aimee Chin, Steven Craig, German Cubas, Elaine Liu, Fan Wang, and the participants of the Applied Microeconomics Workshop at the University of Houston, and European Society for Population Economics (ESPE) for helpful feedback.}}

\maketitle

\begin{center}
\href{https://hhadah.github.io/Attitudes-and-Identity/my_paper/Hadah-Attitudes.pdf}{\textcolor{green!15!black!30!blue}{\footnotesize{\textsc{Click Here to Get the Most Updated Version}}}}
\end{center}

\begin{abstract}
I study the determinants of the choice to identify as Asian among those who could—those whose parents, grandparents, or selves were born in an Asian-speaking country. I find that individuals with Asian ancestry are significantly less likely to self-identify as Asian if they live in states with high levels of implicit ethnic bias. A one standard deviation increase in bias decreases self-reported Asian identity by 7 and 13 percentage points for first and second-generation Asians, respectively. These effects are more prominent among second-generation immigrants with both parents born in an Asian-speaking country than among children of inter-ethnic parents. These findings have implications for the interpretation of research on ethnic gaps in economic outcomes and the correct counting of the population.

\noindent\textbf{Keywords}: Economics of Minorities, Race, and Immigrants; Discrimination and Prejudice

\noindent\textbf{JEL Classification}: I310, J15, J71

\end{abstract}

\vfill
\pagebreak{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main Text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}\label{sec:intro}

The self-reported Hispanic population in the United States has tripled over the last three decades as Hispanics overtook Black American as the largest minority group in America.\footnote{The 2020 Census counted more than 62 million Hispanics---19 percent of the population---triple the number of Hispanics counted three decades earlier \autocite{floodsarahIntegratedPublicUse2021a}. The Hispanic population numbers are based on the author's calculations from the Current Population Survey and US Census data.} An extensive literature on Hispanic--non-Hispanic White gaps has emerged, finding differences in health \autocite{antecolUnhealthyAssimilationWhy2006,antmanEthnicAttritionObserved2016}, wages \autocite{trejoWhyMexicanAmericans1997}, and schooling \autocite{antmanHispanicAmericansLabor2022}. However, defining and measuring these racial and ethnic groups is not straightforward, especially when considering self-reported identity. To the extent that self-reporting Hispanic identity is negatively selected, these gaps could be biased upward.

Various factors, including prejudice, can influence the manner in which individuals select their ethnic identity. In this paper, I explore the determinants of Hispanic identity and how Hispanics self-select into Hispanic and White identities. In particular, I study how bias against minorities influences their decisions to identify, or not, as a member of their ethnic minority. This is important as it affects our interpretations of a variety of findings. First, if individuals react to prejudice by choosing not to identify with their targeted group, standard analyses attempting to identify components of ethnic gaps in outcomes could be overestimated in the most biased states. Second, how individuals identify may impact measured changes in labor market outcomes among groups differentiated by race and ethnicity. As a result, Mexican immigrants' assimilation rates could appear slower than other groups.

I explore how individual characteristics and social attitudes toward Hispanics affect self-reported Hispanic identity. I use identity and ancestry information from the Current Population Survey (CPS) along with a proxy for state-level bias using Harvard's Project Implicit Association Test (IAT) and the American National Election Studies (ANES). \footnote{The IAT data is retrieved from Harvard's Project Implicit \autocite{greenwaldMeasuringIndividualDifferences1998}. The implicit bias toward minorities, as measured by IAT, is widely used by psychologists and is growing in use among economists. IAT scores were shown to be correlated with economic outcomes \autocite{chettyRaceEconomicOpportunity2020,gloverDiscriminationSelfFulfillingProphecy2017}, voting behavior \autocite{friesePredictingVotingBehavior2007}, and health \autocite{leitnerRacialBiasAssociated2016}.} I motivate my analysis with a simple model in the vein of \textcite{akerlofEconomicsIdentity2000}. The model makes an explicit path through which actions affect individuals' utility via their identity and introduces an externality where the actions of others---or prejudice---have different effects on a person's well-being and identity. Therefore, if a person can choose their identity credibly, and this choice is affected by the prejudice of others, then they will choose it to maximize their outcomes.

Measuring identity choices outside of a laboratory is challenging because it requires objective and self-reported identity measures. I use data from a person's birthplace and ancestry to construct an ostensibly objective measure of identity. I find self-reported identity to be negatively correlated with individual and parental characteristics, i.e., parental education. I also find that they are negatively associated with discrimination and ethnic attitudes that reflect the social environment.

Among individuals of Hispanic ancestry, I find that higher state-level bias---against Hispanics---is correlated with a lower self-reported Hispanic identity among Hispanic immigrants. I find that an increase of one standard deviation in bias correlates with a statistically insignificant 2 percentage point decrease in the self-reported Hispanic identity among first-generation immigrants and a statistically significant 4 percentage point decrease among second-generation immigrants. Additionally, a one standard deviation increase in bias is correlated with a 5 percentage points drop in self-reported Hispanic identity among second-generation Hispanic children with both parents born in a Spanish-speaking country. Consequently, as the more economically successful Hispanic immigrants---educated and wealthy immigrants---may not not self-report Hispanic identity, economic research using subjective ethnic measures will overestimate White-Hispanic gaps in the most biased states.

This paper most closely fits in the literature of stratification economics. The interplay between racial identity, economic status, and social outcomes forms a complex web that various scholars have sought to untangle. \textcite{darityEconomicsIdentityOrigin2006,darityPositionPossessionsStratification2022} provide a foundational understanding of the economics of identity and stratification, suggesting that both historical and contemporary economic factors contribute to the persistence of racial norms and inequality. This theme is extended in the context of labor and marital markets by \textcite{goldsmithDarkLightSkin2007,hamiltonSheddingLightMarriage2009,dietteSkinShadeStratification2015} who explore how skin color influences economic and social prospects among African Americans. Similarly, \textcite{golash-bozaLatinoRacialChoices2008} reveal the nuanced racial self-identification processes among Latinos, affected by skin color and discrimination. The significant impact of political and national events on racial identity is also evident in \textcite{masonStigmatizationRacialSelection2014} study of Arab and Islamic Americans post-September 11 and \textcite{masonNotBlackAlone20082017} examination of the 2008 Presidential Election's effect on African American racial identity. These studies collectively underscore the multidimensional nature of racial identity and its profound implications for economic and social stratification. I contribute to the literature of stratification economics by providing evidence that Hispanic identity formation is influenced by societal factors, i.e. discrimination and prejudice.

This paper also fits in the economics of immigration and assimilation. \textcite{abramitzkyCulturalAssimilationAge2016} measured the speed at which immigrants from Europe, Asia, and Latin America assimilate in the United States. They find that assimilation increases over time.\footnote{For more on immigrant assimilation, see \textcite{abramitzkyLeavingEnclaveHistorical2020,abramitzkyIntergenerationalMobilityImmigrants2019,abramitzkyDiscriminationReturnsCultural2020,abramitzkyNationImmigrantsAssimilation2014}} \textcite{foukaImmigrantsAmericansRace2022} investigated the effect of the inflow of Black Americans migrating from the South to the North on the assimilation of European immigrants. The authors found that immigrants in places that received more Black migrants assimilated faster.  \textcite{mengIntermarriageEconomicAssimilation2005} studied the effect of intermarriage on assimilation and found that immigrants who intermarry earn significantly more than those in an endogamous marriage. \textcite{antmanEthnicAttritionObserved2016} show that among immigrants from Mexico, the least economically successful self-identify as being of Mexican origin, while the most successful do not.

Other researchers studied the assimilation of Hispanic immigrants. \textcite{antecolUnhealthyAssimilationWhy2006} documented an interesting puzzle where non-native-born Hispanics have better health outcomes than native-born Hispanics, and \textcite{trejoWhyMexicanAmericans1997} showed that Mexican men earn substantially less than Whites.\footnote{The Hispanic health paradox has led many researchers to try to explain it \autocite{giuntellaAssimilationHealthEvidence2016,giuntellaAccelerationImmigrantUnhealthy2017,giuntellaReasonImmigrationImmigrants2018,giuntellaWhyDoesHealth2017a,antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.} \textcite{smithAssimilationLatinoGenerations2003} offered a more optimistic view of the assimilation of Hispanic immigrants. The longer Hispanic and Latino immigrants reside in the US, the more they can close the educational gap with White men. Moreover, some of the poor showings of how well Hispanic immigrants assimilate in the United States could be explained by ethnic attrition and the use of self-reported Hispanic identity to study Hispanics \autocite{duncanComplexityImmigrantGenerations2017,duncanWhoRemainsMexican2011,mengIntermarriageEconomicAssimilation2005,duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018,antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}. The ethnic attrition was driven by the children of interethnic marriages \autocite{mengIntermarriageEconomicAssimilation2005,duncanEthnicIdentificationIntermarriage2005}. Once the attrition was accounted for, Hispanic immigrants would appear healthier and thus more assimilated than previously thought \autocite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}. This research, however, does not explore whether the self-reported identity itself is a function of other factors like prejudice, which this paper provides evidence on.

This paper is most closely related to \textcite{antmanEthnicAttritionObserved2016,antmanIncentivesIdentifyRacial2015,antmanAmericanIndianCasinos2021} where the authors studied the ethnic attrition of Hispanic immigrants and how minorities change their self-reported identity to changes in policies.\footnote{Ethnic attrition is when a person with Hispanic ancestry fails to self-identify as Hispanic.} Taking into consideration the ethnic attrition that \textcite{antmanEthnicAttritionObserved2016} document, I investigate the determinants of what drives a person to self-report, or not, their Hispanic identity. I aim to decompose some of the complexity associated with endogenous identity by exploring certain personal and environmental determinants of identity. The empirical analysis in this paper documents how certain observable factors, namely personal characteristics and societal attitudes, affect the self-reported identity of Hispanics.

The rest of this paper is structured as follows. First, I will discuss the conceptual framework in section (\ref{sec:model}). Second, I will describe the data I use in section (\ref{sec:data}). Third, I will introduce an empirical model and the results in sections (\ref{sec:empstrat}) and (\ref{sec:results}). Fourth, I will discuss robustness checks and discuss the results in section (\ref{sec:robcheck}). Finally, I conclude in section (\ref{sec:conc}).

\section{Conceptual Framework}\label{sec:model}

I discuss a conceptual framework of identity in the spirit of \textcite{akerlofEconomicsIdentity2000}. A person belongs to some ethnic group, and their actions either affirm or deny their ethnic identity. Actions that deviate from what is proscribed of the ethnic identity are costly.

Formally, a person $i$ belongs to ethnic group $e_i \in \{H, NH\}$, where $H$ is Hispanic and $NH$ is non-Hispanic. Agent $i$'s utility depends on their actions and the extent to which their actions affirm their identity $I_i$:

\begin{equation}
U_i = U_i(\pmb{a_i}, \pmb{a_{-i}}, I_i)\label{eq:util}
\end{equation}

A person's identity, $I_i$, is influenced by their own actions, the actions of others, and the behavior proscribed by their ethnicity. I write this as:

\begin{equation}
I_i = I_i(\pmb{a_i}, \pmb{a_{-i}}; \pmb{B}_{e_{i}})\label{eq:identity}
\end{equation}

Where $\pmb{a_i}$ is the actions of person $i$. $\pmb{a_{-i}}$ is the actions of others that would affect $i$'s identity, i.e., societal bias. $I_i$ is the identity function. Each group has an associated set of behaviors that society proscribes them to conform to, which I denote as $\pmb{B}_{e_{i}}$.\footnote{\textcite{akerlofEconomicsIdentity2000} refer to $B_{e_{i}}$ as proscription.}

A person $i$ chooses action $a_i$ that maximizes their utility function given ethnic group $e_i$, proscribed appropriate behavior $\pmb{B}_{e_{i}}$, and the actions of others $\pmb{a_{-i}}$. This implies the following first-order condition (F.O.C.):

\begin{equation}
\frac{\partial U_i}{\partial a_i} + \frac{\partial U_i}{\partial I_i} \cdot \frac{d I_i}{d a_i} = 0\label{eq:foc}
\end{equation}

Whose solution $a_{i}^{\star}$ yields utility $U_{i}^{\star}$. Now, suppose a person can choose their ethnic identity at a cost of $c$. They will do so if $\tilde{U_{i}}^{\star} \geq  U_{i}^{\star} + c$. Where $\tilde{U_{i}}^{\star}$ is the utility obtained from optimal actions $\tilde{a_i^{\star}}$ under the counterfactual ethnicity.

That is $i$ will change identities when the benefits of doing so $\tilde{U_{i}}^{\star} - U_{i}^{\star}$ exceed the costs $c$. These net benefits are non-zero only if $\frac{d I_i}{d a_i} \neq 0$ and $\frac{\partial U_i}{\partial I_i} \neq 0$. This suggests that an empirical analysis of the determinants of identity choice should focus on: (1) individual characteristics that would lead to different $a_i$ under different identities, (2) contextual characteristics that would lead to different $a_{-i}$---bias---under different identities, (3) the analysis should focus on a sample of the population with small $c$, and (4) the sub-sample with a utility that is greatly affected by their identity---i.e., $\frac{\partial U_i}{\partial I_i} \neq 0$). From the empirical analysis, I could investigate the characteristics that would affect $i$'s actions to take different identities from point (1). These characteristics could be the generation immigrants belong to, whether their parents are interethnic or endogamous, etc. I also investigate how different state-level biases could affect identity. Finally, restricting the sample to people with a small cost of changing identity $c$ guarantees that I do not include populations that would never change identities otherwise---for example, non-Hispanic Whites with non-Hispanic ancestry.

\section{Data}\label{sec:data}

In this section, I describe the datasets I use. To study the association between social attitudes and self-reported Hispanic identity, I must measure subjective and objective Hispanic identities to select a subgroup of Hispanic immigrants for analysis. Thus, I use the Integrated Public Use Microdata Series (IPUMS) Current Population Survey (CPS) \autocite{floodsarahIntegratedPublicUse2021a} and use information on ancestry to construct an objective identity measure. I also use implicit association test (IAT) as a measure of bias, in which bias will shift the costs of identifying as a particular ethnic identity (Hispanic).

\subsection{Measuring Hispanic Identity}\label{subsec:cps}

I measure Hispanic identity using the Current Population Survey (CPS), which allows me to construct an objective measure of the Hispanic identity of minors living with their parents. I will use the information on the place of birth, parent's place of birth, and place of birth of grandparents to construct an objective Hispanic measure.\footnote{Following the works of \textcite{antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}.} Thus, I could perfectly identify and construct a dataset of first-, second-, and third-generation Hispanic immigrants (see Figure \ref{fig:diag} for a visual representation). This will consequently allow me to build an objective measure of the Hispanic identity of minors under the age 17 living with their parents.

The objective measure of identity---unlike the self-reported measure where respondents answer affirmatively when asked if they are Hispanic or Latino---depends on the birthplaces of the individual, their two parents, and four grandparents. Thus, the three identifiable generations are: 1) first-generation immigrants that are born in a Spanish-speaking country with both parents also being born in a Spanish-speaking country, 2) second-generation immigrants are native-born citizens to at least one parent that was born in a Spanish-speaking country, 3) third-generation immigrants are native-born citizens to two native-born parents and at least one grandparent that was born in a Spanish-speaking country.\footnote{I restrict first-generation immigrants whose parents were born in a Spanish country to avoid including naturally born US citizens that were born abroad to US parents.} I restrict the sample to Hispanic Whites, first-, second-, and third-generation immigrants who are 17 year old and younger and still live with their parents between 2004 and 2021. I present a summary of the sample statistics in the Table (\ref{tab:sumstat1}).

The overall sample is 49\% female, and 91\% of the sample self-report their identity as Hispanic---answered yes to the question ``are you Latino/Hispanic?''. The average age is 8.6-year-old. Almost 14\% of mothers have a college degree, and 14\% of fathers have a college degree. I provide the rest of the summary statistics for the overall sample and for both the overall sample and each generation in Table (\ref{tab:sumstat1}).

Moreover, using the place of birth of parents and grandparents, I can objectively identify their ethnic ancestry. Consequently, I can identify different types of parents and grandparents. Using the place of birth of parents, I can divide parents of second-generation children into three objective types: 
\begin{enumerate}
\item Objectively Hispanic-father-Hispanic-mother (HH)
\item Objectively Hispanic-father-White-mother (HW)
\item Objectively White-father-Hispanic-mother (WH)
\end{enumerate}

Similarly, using the place of birth of grandparents, I can divide grandparents of third-generation children into 15 objective types: (1) objectively Hispanic paternal grandfather-Hispanic paternal grandmother-Hispanic maternal grandfather-Hispanic maternal grandmother (HHHH); (2) objectively White paternal grandfather-Hispanic paternal grandmother-Hispanic maternal grandfather-Hispanic maternal grandmother (WHHH); (3) objectively Hispanic paternal grandfather-White paternal grandmother-Hispanic maternal grandfather-Hispanic maternal grandmother (HWHH), etc...

My analysis uses a sub-sample of the US population; I show in Table (\ref{tab:hispbygen}) that I have enough observations in each generation. Consistent with the literature on ethnic attrition among Hispanics, I find significant attrition among third-generation Hispanic immigrants.\footnote{In \textcite{duncanIdentifyingLaterGenerationDescendants2018,duncanSocioeconomicIntegrationImmigrant2018, antmanEthnicAttritionObserved2016,antmanEthnicAttritionAssimilation2020}, the authors find substantial attrition among Hispanics.} These results are displayed in Table (\ref{tab:hispbygen}): most first- and second-generation Hispanic immigrants self-report their identity as Hispanic. Of the first-generation Hispanic immigrants, 96\% self-report their identity as Hispanic. Similarly, 95\% of the second-generation Hispanic immigrants self-report their identity as Hispanic, and 85\% of third-generation Hispanic immigrants identified as Hispanic. That is a more than threefold increase in attrition rates. Most of the attrition among third-generation Hispanics is driven by attrition among the children of interethnic marriages.

\subsection{Measuring Prejudice}

To construct a measure of prejudice, I use the implicit association test and the American National Election Studies (ANES). The implicit association test measures how people associate concepts---for example, Black and dark-skinned people---and evaluations---good, bad. Respondents are asked to quickly match words into categories shown on a screen. Figure (\ref{fig:iatexamples}) shows a few examples of what a test taker would see on a skin tone implicit association test by Harvard's Project Implicit.

I use skin tone implicit association test data to construct a proxy of state-level prejudice \autocite{greenwaldMeasuringIndividualDifferences1998}. This measure has been used in the social sciences, especially in psychology. Previous work has shown that IAT test scores are hard to manipulate \autocite{egloffPredictiveValidityImplicit2002}.

The IAT aims to measure the direction and magnitude of bias in people. It also aims to measure unconscious biases in people or biases that they are unwilling to report. On the one hand, in a meta-analysis of more than 122 papers that used IAT, \textcite{greenwaldMeasuringIndividualDifferences1998} find that IAT measures had significantly higher predictive validity than self-report measures. On the other hand, some research disputes the claims of the IAT's predictive validity.\footnote{Research showed that the IAT tests are correlated with economic outcomes \autocite{chettyRaceEconomicOpportunity2020,gloverDiscriminationSelfFulfillingProphecy2017}, voting behavior \autocite{friesePredictingVotingBehavior2007}, and health \autocite{leitnerRacialBiasAssociated2016}. Participation in the IAT, an online test, is voluntary. Therefore, the samples are not random and might suffer from selection bias in who decides to take the exam. However, bias reflected by IAT scores has been used as a proxy for prejudiced attitudes in an area\autocite{chettyRaceEconomicOpportunity2020}.} The Implicit Association Test (IAT) may not reliably measure or predict implicit prejudice or biased behaviors. Some research shows that implicit biases undergo minor and temporary changes through interventions. Additionally, implicit bias does not predict dictator game giving or being influenced by social pressure, highlighting the distinction between implicit bias and biased actions \autocite{arkesAttributionsImplicitPrejudice2004,forscherMetaanalysisProceduresChange2019,leeDoesImplicitBias2018}. Therefore, I supplement the IAT with a measure of explicit bias from the American National Election Studies (ANES) to construct a composite measure of bias.

I construct another proxy measure of racial animus using the ANES survey \autocite{anes2021} to measure animus, or discrimination, against Black Americans. ANES is a survey that has been conducted since 1948 and is widely used in political science. The survey asks respondents about their attitudes toward different racial groups, voting intentions, and other political questions. I use several questions from the ANES surveys conducted between 2004 and 2020 to measure racial animus. The racial animus index is constructed by taking the average of the responses to several questions measuring racial animus. \footnote{The questions used are similar to those used by \textcite{charlesPrejudiceWagesEmpirical2008}. The questions are: (1) ``Conditions Make it Difficult for Blacks to Succeed'', (2) ``Blacks Should Not Have Special Favors to Succeed'', (3) ``Blacks Must Try Harder to Succeed'', (4) ``Blacks Gotten Less than They Deserve Over the Past Few Years'', and (5) ``Feeling Thermometer Toward Hispanics.''}

Moreover, to reduce attenuation bias and measurement error, I follow \textcite{lubotskyInterpretationRegressionsMultiple2006} in constructing a composite bias measure using both the IAT and the ANES racial animus measure.\footnote{More on the method in the Data Online Appendix, see Section~\ref{sub:lw-bias}.} Figure (\ref{fig:skiniat}) shows a graphical representation of the bias measure over time in the most and least biased locations. Figure (\ref{fig:hispanic-twostates}) shows a graphical representation of self-reported Hispanic identity in the two most and least biased locations. A lower score implies less bias, whereas a higher score implies higher racial animus. A one standard deviation increase in bias is equivalent to moving from Washington, DC, or Vermont to North Dakota in 2020. I also show the state-level average bias over time in the maps in Figure (\ref{fig:skiniat-maps}) and the overall average from 2004 to 2021 in Figure (\ref{fig:iat-map-all}).
\section{Estimation and Results}\label{sec:empstrat}

To understand the association between Hispanic self-identity and state-level bias, I estimate regressions of the following form for each generation $g$:

\begin{align}
H_{ist}^g &= \beta_1^g Bias_{st} + \beta_2^g DadCollegeGrad_{ist} + \beta_3^g MomCollegeGrad_{ist} \nonumber \\ 
            &+ \beta_4^g Women_{ist} + X_{ist}^g\pi + \gamma_{rt} 
           + \varepsilon_{ist}; 
           \text{where } g \in \{1,2,3\} \label{eq:identity_reg_bias}
\end{align}

Where $H_{ist}^g$ be the self-reported Hispanic identity of person $i$ in state $s$ at the time of interview $t$, let $Bias_{st}$ be the average state-level bias in state $s$ at time $t$, $DadCollegeGrad_{ist}$, and $MomCollegeGrad_{ist}$ are indicator variables that are equal to one if the father or mother graduated from college, $Women_{ist}$ is an indicator variable for sex, and $X_{ist}$ is a vector of controls.\footnote{The controls include quartic age, fraction of population that is Hispanic in state $s$, type of parents (WH, HW, or HH), type of grandparents (HHHH, HHHW, etc.), and dummy variables the generation to which person $i$ belong.} Additionally, $\gamma_{rt}$ is region-time fixed effects that controls for region $\times$ year specific shocks.\footnote{I do not include state fixed effects because of lack of with-in state variation in bias.} The region $\times$ year also controls for systematic differences between regions in the overall Hispanic population and bias toward Hispanics, even if they vary over time. Throughout the analysis, I cluster the standard errors at the state level to account for correlation in the error term $\varepsilon_{ist}$ within a state, overtime.

Since the specification includes region $\times$ year, $\gamma_{rt}$, the $\beta_1^g$ coefficient summarizes individual's $i$ responsiveness to state-level bias changes in the state which they live. In other words, $\beta_1^g$ captures the association between self-reported Hispanic identity and state-level bias across states within a Census division region. Additionally, the $\gamma_{rt}$ fixed effects account for any regional and national trends in bias over time. Consequently, $\beta_1^g$ provide the correlation between self-reported Hispanic identity and state-level bias above and beyond the national and regional trends in bias. If individuals in states within a region responded similarly to changes in state-level bias, then $\beta_1^g$ will be equal to zero.

\section{Results}\label{sec:results}
The results from the regression framework described above provide consistent evidence aligning with the following findings. First, state-level bias is negatively associated with self-reported Hispanic identity. Second, first- and second-generation Hispanic immigrant children of endogamous marriages' self-reported Hispanic identity are more negatively associated with state-level bias.

I report the main results of estimating equation (\ref{eq:identity_reg_bias}) in Figure (\ref{plot01-regression-gen}). I present the results of estimating the main specification for second-generation immigrants in panel (A) and for sub-samples of HH, HW, and WH children in panels (B), (C), and (D), respectively. I find that bias and self-reported Hispanic identity are negatively associated, while parental education and self-reported Hispanic identity are positively associated. A one standard deviation increase in state-level bias is associated with a 4 percentage points decrease in self-reported Hispanic identity. Among first- and second-generation Hispanic immigrants, a one standard deviation increase in state-level bias is associated with 2 and 4 percentage points decrease in self-reported Hispanic identity. The coefficient is not statistically significant for first-generation, but the confidence interval is mostly in the negative territory.

I report the results of the same regression but on sub-samples of second-generation immigrants by type of parents---interethnic and endogamous parents---in Figure (\ref{plot01-regression-byparent}). I present the results of estimating the main specification on second-generation immigrants in panel (A) and on sub-samples of HH, HW, and WH children in panels (B), (C), and (D), respectively. I find that children of endogamous Hispanic marriages are more influenced by state-level bias than those of interethnic marriages. I find that a one standard deviation increase in state-level bias is associated with 5 percentage points decrease in self-reported Hispanic identity among children of HH parents.

I also report the results of the regression on sub-samples of third-generation immigrants by the number of Hispanic grandparents in Table (\ref{regtab-bygrandparents}). I find that state-level bias is not significantly associated with the self-reported Hispanic identity of children of interethnic grandparents. However, it is negatively associated with lower self-reported Hispanic identity of children of Hispanic grandparents. I find that a one standard deviation increase in state-level bias is associated with a 6 percentage points decrease in self-reported Hispanic identity.

\section{Robustness Checks and Discussions} % (fold)
\label{sec:robcheck}

In this section, I explore the empirical relationship between state-level bias and interethnic marriages, as well as the migration patterns of second-generation Hispanic immigrants as robustness checks to my main analysis and the effect of proxy response on my results. I examine the impact of state-level biases on the likelihood of interethnic marriages, focusing on interethnic couples, and the migration decisions of Hispanic individuals within the United States.

I investigate the relationship between state-level bias and interethnic marriages. To this purpose, the regression specifications for the estimation will be as follows:

\begin{align}
Interethnic_{ist}^2 &= \beta_1^2 Bias_{st} + X_{ist}^2\pi + \gamma_{rt} 
            + \varepsilon_{ist}  \label{eq:inter-interethnic} 
\end{align}

Where $Interethnic_{ist}^2$ is an indicator variable that is equal to one if a couple is interethnic, i.e., a Hispanic husband-White wife or a White husband-Hispanic wife. $Bias_{st}$ is the average bias in state $s$ at time $t$, and $X_{ist}^2$ is a vector of partner-specific controls that would affect a marriage match that includes the wife's and husband's education, age, and years since immigrating to the United States.

I present the results of estimating equation (\ref{eq:inter-interethnic}) in Table (\ref{regtab-logit-02}). I find that a one standard deviation increase in bias increases the probability of having interethnic parents by 3 percentage points. Moreover, I break down the analysis by the ethnicity of the couples. A one standard deviation increase in state-level bias is associated a 3 percentage points increase in the chances of a Hispanic husband marrying a White wife. A one standard deviation increase in state-level bias is associated a 4 percentage points increase in the chances of a Hispanic wife having a White husband. The fact that bias and interethnic marriage are positively correlated could be a result of the fact that Hispanic immigrants in states with high bias might aim to decrease the likelihood that their children will display Hispanic ethnicity signals. For example, Hispanic women in high bias states might marry a non-Hispanic White husband, so their children will have a non-Hispanic last name.

I am also interested in investigating the relationship between state-level bias and migration. As the CPS does not report a person's birth state, I use the 2004-2021 Censuses to construct a sample of second-generation Hispanic immigrants \autocite{floodsarahIntegratedPublicUse2021}. I construct a mover variable to indicate whether these second-generation Hispanic immigrants have moved from their birth state to another state. For this purpose, I use the following models to estimate the relationship between state-level bias and migration:

\begin{align}
BirthPlaceMigration_{ist}^2 &= \beta_1^2 Bias_{st} 
                   + X_{ist}^2\pi + \gamma_{rt} 
                   + \varepsilon_{ist} \label{eq:migration-3} \\
BirthPlaceMigration_{ilb}^2 &= \beta_1^2 Bias_{lb} 
                   + X_{ilb}^2\pi + \gamma_{lb} 
                   + \varepsilon_{ilb} \label{eq:migration-4}
\end{align}

Where $BirthPlaceMigration_{ist}^2$ is an indicator variable equal to one if person $i$ in state $s$ at the interview $t$ lives in a state that is different from his or her birth state and zero otherwise. $BirthPlaceMigration_{ilb}^2$ is an indicator variable that is equal to one if person $i$ in birthplace $l$ does not currently live in the same state he or she lived in at the year of birth $b$ and zero otherwise. The analysis, restricted to second-generation Hispanic immigrants with both parents born in a Spanish-speaking country, uses equations (\ref{eq:migration-3}) and (\ref{eq:migration-4}).

Furthermore, I use two ways to define the bias variable to study the relationship between bias and the migration variables introduced above. In the first specification from equation (\ref{eq:migration-3}), I estimate the relationship between the average bias at the time of the interview $t$ in state $s$ and $BirthPlaceMigration_{ist}^2$. In the second specification from equation (\ref{eq:migration-4}), I estimate the relationship between the average bias in birth state $l$ at the year of birth $b$ and $BirthPlaceMigration_{ilb}^2$.

I also estimate whether those who self-identify as Hispanic tend to move from high-bias to low-bias states. The estimation equation for the relationship is:

\begin{align}
Y_{ist} &= \beta_0 + \beta_1^2 Hispanic_{ist} +
                   X_{ist}^2\pi
                   + \varepsilon_{ist} \label{eq:migration-5}
\end{align}

Where $Y_{ist} \equiv Bias_{ist} -  Bias_{ilb}$, $Bias_{ist}$ is $i$'s state-level bias in state $s$ at the time of interview $t$, and  $Bias_{ilb}$ is $i$'s state-level bias in birth state $l$ at the birth year $b$. The analysis is restricted to second-generation Hispanic immigrants with both parents born in a Spanish-speaking country who migrated from the state they were born in $b$ to another state $s$.

The results of estimating equations (\ref{eq:migration-3}), (\ref{eq:migration-4}), and (\ref{eq:migration-5}) are shown in Table (\ref{regtab-mig-01}) in columns (1), (2), and (3) respectively. I find that among second-generation immigrants, there is no significant correlation between bias and migration decisions. Among second-generation Hispanic immigrant movers, those who self-report Hispanic identity live in states with 0.06 standard deviations more biased than the state where they were born. Even though this result shows that there is selection into more biased states among second-generation immigrants, it does not affect my main results showing a correlation between bias and self-reported Hispanic identity. Since those identifying as Hispanics are the movers, my assessments of the relationship between bias and self-reported Hispanic identity might underestimate the effect of bias.

The findings presented in this paper indicate a negative correlation between bias and the self-reported Hispanic identity among Hispanic immigrants. While my aim is not to establish a causal effect of bias on self-reported Hispanic identity, I intend to illustrate a correlation between bias and self-reported identity. This correlation suggests that depending on the levels of bias in a state, racial and ethnic gaps that rely on self-reported identity might either overestimate or underestimate the effect of discrimination.

There are a couple of concerns with this analysis. First, the self-reported identity in the Current Population Survey (CPS) is reported by a household respondent—parent or adult caregiver. Thus, the 'self-reported' ethnic identity might not reflect a child's true identity. I view the identity that a parent or a caregiver reports as an accurate representation of the child's identity since parents are essential in shaping their children's sense of self. Also, I compare states with a high and low bias for my analysis. The estimates will not be threatened if the likelihood of self-reporting does not differ between these states.

Moreover, \textcite{duncanIntermarriageIntergenerationalTransmission2011} show that reported Hispanic identification does not vary with who is the household respondent. Additionally, I present the main effect of self-reported Hispanic identity by the household respondent in Table (\ref{tab:hispbyproxy}) and the results to the estimation of equation (\ref{eq:identity_reg_bias}) by the proxy respondent in Table (\ref{regtab-proxy-01}) across all generations. The main effect of the reported Hispanic identity of children is 93 percentage points when the mother is the proxy, 92 percentage points when the father is the proxy, and 96 percentage points when the child or another caregiver was the household respondent.\footnote{According to the Current Population Survey (CPS), a person can be the household respondent if they are at least 15 years old and have enough knowledge about the household. Thus, when the proxy is 'self,' the respondent is between the ages of 15 and 17.} The estimation of equation (\ref{eq:identity_reg_bias}) by the proxy respondent, Table (\ref{regtab-proxy-01}), mostly yields a negative effect of bias on self-reported Hispanic identity for all types of proxy respondents.

A second concern is that the IAT is voluntary and not representative of the population. While I do not claim that the IAT  as a proxy for bias will represent the population, \textcite{egloffPredictiveValidityImplicit2002} show that they are hard to manipulate. Several studies have shown that IAT is correlated with economic outcomes \autocite{chettyRaceEconomicOpportunity2020,gloverDiscriminationSelfFulfillingProphecy2017}, voting behavior \autocite{friesePredictingVotingBehavior2007}, decision-making \autocite{bertrandImplicitDiscrimination2005,carlanaImplicitStereotypesEvidence2019}, and health \autocite{leitnerRacialBiasAssociated2016}. Another concern could be that the IAT test takers' characteristics change over time and, thus, are not the same. I address this concern by including non-parametric region $\times$ year fixed effects that would control for the systematic difference in the characteristics of test takers between regions. These changes will be controlled for as long as the differences in the characteristics between test takers do not vary across states within a region. Most importantly, I use the ANES racial animus measure to construct a composite measure of bias that reduces measurement error using \textcite{lubotskyInterpretationRegressionsMultiple2006}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendix A, Solution and Estimation Details
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Set equation, figure, table indexing
\renewcommand{\thefigure}{A.\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{A.\arabic{table}}
\setcounter{table}{0}
\renewcommand{\theequation}{A.\arabic{equation}}
\setcounter{equation}{0}
\renewcommand{\thefootnote}{A.\arabic{footnote}}
\setcounter{footnote}{0}
